---
title: "Gonzales_Lab2"
author: "Elise Gonzales, Collaborator: Amrit Sandhu"
date: "2023-01-24"
output: pdf_document
---

Source Lab 1

```{r}
# Source from lab 1
sourceDir <- "/Users/elise/Documents/MEDS/Courses/EDS232/LABS/eds-232-machine-learning/Lab1.Rmd"
library(knitr)
source(knitr::purl(sourceDir, quiet=TRUE))
```

Today we will be continuing the pumpkin case study from last week. We will be using the data that you cleaned and split last time (pumpkins_train) and will be comparing our results today to those you have already obtained, so open and run your Lab 1 .Rmd as a first step so those objects are available in your Environment (unless you created an R Project last time, in which case, kudos to you!).

Once you have done that, we'll start today's lab by specifying a recipe for a polynomial model. First we specify a recipe that identifies our variables and data, converts package to a numerical form, and then add a polynomial effect with step_poly()

```{r}
# Specify a recipe
poly_pumpkins_recipe <-
  recipe(price ~ package, data = pumpkins_train)  |> 
  step_integer(all_predictors(), zero_based = TRUE)  |>  
  step_poly(all_predictors(), degree = 2)
```

How did that work? Choose another value for degree if you need to. Later we will learn about model tuning that will let us do things like find the optimal value for degree. For now, we'd like to have a flexible model, so find the highest value for degree that is consistent with our data.

Polynomial regression is still linear regression, so our model specification looks similar to before.

```{r}
# Create a model specification called poly_spec
poly_spec <- linear_reg()  |>  
  set_engine("lm")  |>  
  set_mode("regression")
```

Question 1: Now take the recipe and model specification that just created and bundle them into a workflow called poly_df.

```{r}
# Bundle recipe and model spec into a workflow
poly_wf <- workflow() |> 
  add_recipe(poly_pumpkins_recipe) |> 
  add_model(poly_spec)
  
```

Question 2: fit a model to the pumpkins_train data using your workflow and assign it to poly_wf_fit

```{r}
# Create a model
poly_wf_fit <- poly_wf |> 
  fit(data = pumpkins_train)
  
```

```{r}
# Print learned model coefficients
poly_wf_fit
```

```{r}
# Make price predictions on test data
poly_results <- poly_wf_fit  |>  
  predict(new_data = pumpkins_test)  |>  
  bind_cols(pumpkins_test  |>  
              select(c(package, price)))  |>  
  relocate(.pred, .after = last_col())

# Print the results
poly_results  |>  
  slice_head(n = 10)

```

Now let's evaluate how the model performed on the test_set using yardstick::metrics().

```{r}
poly_met <- metrics(data = poly_results, 
                    truth = price, 
                    estimate = .pred)
poly_met
```

Question 3: How do the performance metrics differ between the linear model from last week and the polynomial model we fit today? Which model performs better on predicting the price of different packages of pumpkins?

*RMSE: "Root Mean Square Error"; A lower value is better than a higher one.The polynomial (4.72) is more accurate than the linear result (7.23).*

*RSQ: "Root Squared"; 0 means no-fit and 1 means perfect fit. The polynomial result (0.777) has a better fit compared to the linear model (0.495).*

*The polynomial model performs better than the linear regression.*

Let's visualize our model results. First prep the results by binding the encoded package variable to them.

```{r}
# Bind encoded package column to the results
poly_results <- poly_results  |>  
  bind_cols(package_encode  |>  
              rename(package_integer = package))  |>  
  relocate(package_integer, .after = package)


# Print new results data frame
poly_results  |>  
  slice_head(n = 5)
```

OK, now let's take a look!

Question 4: Create a scatter plot that takes the poly_results and plots package vs. price. Then draw a line showing our model's predicted values (.pred). Hint: you'll need separate geoms for the data points and the prediction line.

```{r}
# Make a scatter plot
poly_results  |>  ggplot(aes(x=package_integer)) +
  geom_point(aes(y = price), color = "navy") +
  geom_point(aes(y=.pred), color = "orange") +
  geom_line(aes(y=.pred), color = "gray") +
  labs(title = " Package vs. Price",
       x = "Package",
       y = "Price")

```

You can see that a curved line fits your data much better.

Question 5: Now make a smoother line by using geom_smooth instead of geom_line and passing it a polynomial formula like this: geom_smooth(method = lm, formula = y \~ poly(x, degree = 3), color = "midnightblue", size = 1.2, se = FALSE)

```{r}
# Make a smoother scatter plot 
poly_results  |>  ggplot(aes(x= package_integer, y = price)) +
  geom_point(aes(y = price), color = "navy") +
  geom_point(aes(y = .pred), color = "orange") +
  geom_smooth(method = lm,
              formula = y ~ poly(x, degree = 3),
              color = "gray",
              size = 1.2,
              se= FALSE)  +
   labs(title = " Package vs. Price",
       x = "Package",
       y = "Price")
```

OK, now it's your turn to go through the process one more time.

Additional assignment components :

6\. Choose a new predictor variable (anything not involving package type) in this data set.

```{r}
# Name predictor variables
names(baked_pumpkins)
```

My new predictor variable is **variety**.

7\. Determine its correlation with the outcome variable (price). (Remember we calculated a correlation matrix last week)

```{r}
# Find the correlation between the variety and the price
cor(baked_pumpkins$variety, baked_pumpkins$price)
```

```{r}
# Obtain correlation matrix and drop low_price and high
corr_matrix <- cor(baked_pumpkins  |>  
  select(-c(low_price, high_price, month, day, date)))

# Make a correlation plot between the variables
corrplot(corr_matrix, method = "shade", shade.col = NA, tl.col = "black", tl.srt = 45, addCoef.col = "black", cl.pos = "n", order = "original")
```

8.  Create and test a model for your new predictor:

-   Create a recipe

```{r}
var_recipe <- recipe(data = pumpkins_train, price ~variety) |> 
  step_integer(all_predictors(), zero_based = TRUE)
```

-   Build a model specification (linear or polynomial)

```{r}

var_poly_spec <- linear_reg()  |>  
  set_engine("lm")  |>  
  set_mode("regression")

```

-   Bundle the recipe and model specification into a workflow

```{r}
var_poly_wf <- workflow() |>
  add_recipe(var_recipe) |>
  add_model(var_poly_spec)

```

-   Create a model by fitting the workflow

```{r}
var_fit <- poly_wf  |>  
  fit(data = pumpkins_train)
```

-   Evaluate model performance on the test data

```{r}
var_results <- pumpkins_test  |>  
  select(c(variety, price))  |> 
  bind_cols(predictions)

head(var_results)
```

-   Create a visualization of model performance

```{r}
#Encode package column
var_encode <- var_recipe |>  
  prep() |>  
  bake(new_data = pumpkins_test) |>  
  select(variety)

#Bind encoded package column to the results
 plot_results_var <- results_var %>%
 bind_cols(var_encode %>%
               rename(var_integer = variety)) %>%
  relocate(var_integer, .after = variety)
 
#Print new results
plot_results |> 
  slice_head(n = 5)

#Scatter plot
plot_results_var |> 
  ggplot(mapping = aes(x = var_integer, y = price)) +
   geom_point(size = 1.6, alpha = 0.2) +
   geom_line(aes(y = .pred), color = "gray", size = 1.2) +
   labs(x = "Variety", y = "Price", title = "Variety vs. Price")

# Graph with geom_smooth
plot_results_var  |>
  ggplot(mapping = aes(x = variety, y = price)) +
   geom_point(size = 1.6, alpha = 0.2, color = "navy") +
    geom_smooth(method = lm, formula = y ~ poly(x, degree = 2), color = "orange", size = 1.2, se = FALSE) + labs(x = "Variety", y = "Price", title = "Variety vs. Price")

```
